### 2014-01-28 thesis/src/chapters/chapter-design.tex
Thesis: added arrows to figure

Service consumers can be both end-users or other system interfaces, Cloud-SAP is suitable for both cases. Cloud-SAP is entirely cloud provider and resources agnostic as long as they externalise management interfaces.

Successive sections define core layers of a Cloud-SAP and give insights into possible implementation decisions, while the next chapter presents proof-of-concept implementation.

@@

\subsection{Orchestrating autonomic manager}

@@

% radek
\section{Autonomic cloud federation manager}

>>>>

Application providers can be both end-users or other system interfaces, Cloud-SAP is suitable for both cases. Cloud-SAP is entirely cloud providers and resources agnostic as long as they externalise management interfaces.

Successive sections define core layers of a Cloud-SAP and give insight into possible implementation decisions, while the next chapter presents a proof-of-concept implementation.

@@

\subsection{Orchestrating autonomic manager}
While an autonomic manager which manages a single resource may be satisfactory enough in many business cases, there are situations when additional coordination among managers is essential to attain the certain goal. This can result in incorporating autonomic behaviour within an entity in a system-wide scope.

Such a coordination can be achieved by introducing another autonomic manager whose sole purpose is to harmonize the work of dependent ones by extensive use of their \emph{sensor} and \emph{effector} interfaces. These managers are called \emph{orchestrating managers}.

When it comes to the complex task of ensuring the Quality of Service requirements for a client of a cloud computing environment, it is vital to ensure proper mechanisms that would allow seamless cooperation among different cloud providers. Such cooperation could result in a migration of an application between different clouds, market-driven choice of the cloud providers and so on.

As the exemplary actions can be regarded complex ones, they cannot be associated with actions taken by autonomic managers of only one type. Thus, an orchestrating autonomic manager should coordinate other managers of any type -- an arbitrary mixture of self-healing, self-configuring, self-optimizing and self-protecting managers. Employment of these concepts in Cloud-SAP will be discussed later in detail.

@@

% radek
\section{Autonomic cloud federation manager}
\emph{Autonomic cloud federation manager} is a building block of the proposed architecture that forms the highest layer of its model (see figure \ref{img:csap-layered-structure}). Additionally, it is an element which directly cooperates with a client by capturing and handling their requests.

This manager has two main responsibilities, which include
\begin{inparaenum}[1)]
\item handling deployment of a service requests from the clients and subsequent requests directly applied to that service (e.g. queries about its state), 
\item scaling the previously-deployed service across different cloud vendors.
\end{inparaenum}
The former involves capturing the request from a client, determining the best cloud vendor or vendors for the described service and its deployment. This behaviour is strictly related with \emph{plan} and \emph{execute} functions of an autonomic manager. The latter, on the other hand, involves \emph{monitoring} the dependant autonomic managers that manage in an autonomic fashion cloud instances, \emph{analysing} obtained knowledge and deciding if taking any auto-scaling action is actually necessary to ensure satisfying Quality-of-Service requirements. If such an action is needed, it must be \emph{planned} and \emph{executed}.

The use case of scaling a service withing a federation of cloud vendors can be considered an action which coordinates cloud instance managers. It is perfectly legitimate to view the cloud federation manager as an \emph{orchestrating} one. What is more, such a coordination requires cooperation with not only autonomic managers of a single type, but with the ones that are a mixture of self-healing, self-configuring, self-optimizing and self-protecting. Taking this into account, the orchestration can be classified as an \emph{orchestration across disciplines} \cite{IBM06}.

\subsection{Structure}
It is clear from the description of requirements of this autonomic manager in the previous section that \emph{plan} and \emph{execution} functions are common for both of them. Therefore the \emph{cloud federation manager} is composed of two internal autonomic managers, which perform complete different functions, the first one -- \emph{monitor} and \emph{analyse} and the second -- \emph{plan} and \emph{execute}, yet they cooperate with each other to realize the full control loop.

To be able to handle incoming requests, the manager exposes API a client communicates with. The complete structure of this component is shown in figure \ref{img:orchestrating-autonomic-manager}. In the next sections a detail description of each internal autonomic manager will be presented.

\begin{figure}[!ht]
  \begin{center}
    \includegraphics[width=0.7\textwidth]{chapter-design/csap-orchestrating-autonomic-manager}
  \end{center}
  \caption{Design of an orchestrating autonomic manager}
  \label{img:orchestrating-autonomic-manager}
\end{figure}

\subsection{Managed resource}
As it was stated in the previous sections and can be seen in figure \ref{img:orchestrating-autonomic-manager}, \emph{cloud instance autonomic managers} are managed resources of this manager. Each cloud instance is a deployment environment for a service or a stack, i.e. it is possible to span the service across different clouds by deploying stacks which make up the given service to different providers. Of course it is legitimate to deploy the whole service to only one provider.
\subsubsection*{Touchpoints}
As the cloud federation manager has to gain information about the state of a given cloud instance and be able to order the deployment of a stack or a service, the cloud instance exposes its manageability interface and its two types:
\begin{asparaenum}
\item[\textbf{Sensors}] -- set of attributes which forms the state of a given cloud instance. Attributes should include inter alia overall resources usage, pricing, geographical location. The more thorough elaboration will be held in the section devoted to \emph{plan and execution manager}.
\item[\textbf{Effectors}] -- set of stack/service management operations, such as deploy, migrate or delete.
\end{asparaenum}

\subsection{Plan and Execution Autonomic Manager}
\subsection{Monitor and Analyze Autonomic Manager}
